---
title: ""
author: "Laurie Lyon"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse) 
library(ape)
library(rstatix)
library(ggpubr)
library(ggbeeswarm)
```
Read in files 
```{r}
all_ints <- read.csv("../data/20250925_M01_all_interactions_90pct_1000.csv")
agora2_genus_fp <- "~/Projects/micom-time-series/data/models/agora201_refseq216_genus_1/a4ce3649-c8b8-4893-8bd2-4886aa1f40e8/data/"
```

Generate list of all genera (removing .json from each file in teh AGORA2 genus reference database)
```{r}
agora2_filename_list <- list.files(agora2_genus_fp)

clean_genus_list <- c()

for (genus in unlist(agora2_filename_list)) {
  genus_only <- gsub(x = genus, pattern = ".json", replacement = "")
  clean_genus_list <- append(clean_genus_list, genus_only)
}
```

Create a genus_j and genus_i column to be matched with cleaned AGORA2 genus names
```{r}
#use regex to remove everything before the genus name from the taxonomy
#and everything after the genus name (;s__)
genus_filtered <- all_ints %>% 
  mutate(Genus_j = str_extract(Taxon_j, "(?<=g__).*"),
         Genus_j = str_extract(Genus_j, ".*?(?=;s__)"), 
         Genus_i = str_extract(Taxon_i, "(?<=g__).*"),
         Genus_i = str_extract(Genus_i, ".*?(?=;s__)") 
  ) %>% 
  #filter out NAs (no genus level taxonomy)
  filter(!is.na(Genus_j), !is.na(Genus_i))

#filter for pairs where both genera are found in the AGORA database
AGORA_filtered <- genus_filtered %>% 
  #making column with 'present' if that genus list is in the AGORA cleaned genus list
  #or absent if it is not found
  mutate(AGORA_j = ifelse(Genus_j %in% clean_genus_list, 'present', 'absent'),
         AGORA_i = ifelse(Genus_i %in% clean_genus_list, 'present', 'absent')
  ) %>% 
  #filter for pairs where both are present 
  filter(AGORA_j == 'present' & AGORA_i == 'present',
         #and when the genus of OTUj and OTUi aren't equal
         Genus_j != Genus_i)

#seeing what it looks like if we leave all the instances where Genusi = Genusj
AGORA_filtered_w_ijmatch<- genus_filtered %>% 
  #making column with 'present' if that genus list is in the AGORA cleaned genus list
  #or absent if it is not found
  mutate(AGORA_j = ifelse(Genus_j %in% clean_genus_list, 'present', 'absent'),
         AGORA_i = ifelse(Genus_i %in% clean_genus_list, 'present', 'absent')
  ) %>% 
  #filter for pairs where both are present 
  filter(AGORA_j == 'present' & AGORA_i == 'present')
```

*Stats*

```{r}
#with rstatix
full_filter_stats_df <- AGORA_filtered %>%
  #pairwise wilcox rank test (nonparametric t-test)
  wilcox_test(i_j_dist ~ interaction) %>%
  #adjust for multiple tests by benjamini hochberg
  adjust_pvalue(method = "BH") %>%
  #add the column for significance stars
  add_significance("p.adj") 

ijmatch_filter_stats_df <- AGORA_filtered %>%
  #pairwise wilcox rank test (nonparametric t-test)
  wilcox_test(i_j_dist ~ interaction) %>%
  #adjust for multiple tests by benjamini hochberg
  adjust_pvalue(method = "BH") %>%
  #add the column for significance stars
  add_significance("p.adj")
```

Plot code pasted from interactions_phylogeny but this time only on pairs within AGORA
```{r, fig.width = 4, fig.height= 4}
(agora_tlc_phylo_dist_plot <- AGORA_filtered %>% 
   ggplot(aes(x = interaction, y = i_j_dist)) +
   #geom_violin() + 
   geom_boxplot() +
   geom_quasirandom(alpha = 0.3) +
   theme_bw() +
   labs(title = "Phylogenetic Distance by Interaction Type", 
        x = "Interaction Type", 
        y= "Phylogenetic Distance Between OTUj and OTUi") +
   stat_pvalue_manual(full_filter_stats_df,
     label = "p.adj.signif",
     tip.length = 0.01,
     label.size = 4,
     y.position = c(3.2, 3.5, 3.8))
)

(agora_tlc_phylo_dist_plot_ijmatch <- AGORA_filtered_w_ijmatch %>% 
   ggplot(aes(x = interaction, y = i_j_dist)) +
   #geom_violin() + 
   geom_boxplot() +
   geom_quasirandom(alpha = 0.3) +
   theme_bw() +
   labs(title = "Phylogenetic Distance by Interaction Type", 
        x = "Interaction Type", 
        y= "Phylogenetic Distance Between OTUj and OTUi") +
   stat_pvalue_manual(ijmatch_filter_stats_df,
     label = "p.adj.signif",
     tip.length = 0.01,
     label.size = 4,
     y.position = c(3.2, 3.5, 3.8))
)
```

There are multiple OTUs per genera, so I want to find the number of unique combinations of interaction type (+/-/NULL), Genus_j, and Genus_i)
```{r}
unique_genus_int_combos <- AGORA_filtered %>% 
  select(interaction, Genus_j, Genus_i) %>% 
  distinct()

unique_genus_combos <- AGORA_filtered %>% 
  select(Genus_j, Genus_i) %>% 
  distinct()
```
