---
title: "lag_regress_permutation_checks"
author: "Laurie Lyon"
date: "2025-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(cowplot) # to plot grid
library(glue) # to format and interpolate a string
library(lme4) # for linear modeling 
library(tidyverse) 
library(ggh4x) # ggplot2 extension, includes nested facets
library(broom)
library(dplyr)
library(ggplot2)
library(ggpubr)
```

Load in the docs I have from Casey of his significant pairwise interactions and my current lagged regression outputs 
```{r}
casey_sig_ints <- read.csv("../../interactions_taxonomy_files_to_merge/best_fit_organized_interactions.csv")

example_output <- read.csv("../results/lagged_regression/251003_M01_90pct_1000.csv")
#example_output <- read.csv("../results/lagged_regression/M01_updated_perm_queue_0709_1M_1.csv")
example_output1 <- read.csv("../results/lagged_regression/M01_updated_perm_queue_0709_100k_1.csv")
#example_output_old <- read.csv("../results/lagged_regression/M01_perm_test_0618_short_100k_1.csv")

```

```{r}
sum(example_output$pseudo_pval_shuffled_fdr <= 0.1)
sum(example_output$pseudo_pval_shuffled < 0.05)
sum(example_output$pseudo_pval_iaaft_fdr <= 0.05)
sum(example_output$pseudo_pval_shuffled <= 0.1 & example_output$pseudo_pval_iaaft_fdr <= 0.1)
sum(example_output$pseudo_pval_shuffled_fdr <= 0.1 & example_output$pseudo_pval_iaaft_fdr <= 0.1)
```
```{r}
sum(example_output$pseudo_pval_iaaft_fdr <= 0.05)
sum(example_output1$pseudo_pval_iaaft_fdr <= 0.05)
sum(example_output$pseudo_pval_iaaft_fdr <= 0.05 & example_output1$pseudo_pval_iaaft_fdr <= 0.05)
```
```{r}
supp_table <- read.csv("~/Downloads/Bangladesh_supplemental_table_1.csv")
```

```{r}
#merge ones that Casey found to be significant vs. all of my outputs. Both OTU j and OTU i need to match
inner_join_merged_df <- inner_join(casey_sig_ints, example_output, by = c("Feature_ID_j" = "j", "Feature_ID_i" = "i"))

#checking which were found in casey's sig ints but not found in my output
not_in_my_data <- anti_join(casey_sig_ints, example_output, by = c("Feature_ID_j" = "j", "Feature_ID_i" = "i"))
#not in my data makes sense - all come from F01 or M02

colnames(inner_join_merged_df)

#selected_df <- inner_join_merged_df[, c("Feature_ID_j", "Feature_ID_i", "r_coef", "p.value", "r_obs", "p_time", "mean_time", "p_val", "mean_val", "p_amp", "mean_amp", "p_spec", "mean_spec", "p_time_fdr", "p_val_fdr", "p_amp_fdr", "p_spec_fdr", "p_time_reject", "p_val_reject", "p_amp_reject", "p_spec_reject")]

#commented out 10-3-25 because I commented out shuffling to speed up lagregress script
#selected_df <- inner_join_merged_df[, c("Feature_ID_j", "Feature_ID_i", "r_coef", "p.value", "spear_rho_real", "mean_rho_shuffled", "pseudo_pval_shuffled", "mean_rho_iaaft", "pseudo_pval_iaaft", "pseudo_pval_shuffled_fdr", "pseudo_pval_shuffled_signif", "pseudo_pval_iaaft_fdr", "pseudo_pval_iaaft_signif")]

selected_df <- inner_join_merged_df[, c("Feature_ID_j", "Feature_ID_i", "r_coef", "p.value", "spear_rho_real", "mean_rho_iaaft", "pseudo_pval_iaaft", "pseudo_pval_iaaft_fdr", "pseudo_pval_iaaft_signif")]

#selected_df$sig_p_time <- selected_df$p_time < 0.05
#selected_df$sig_p_val <- selected_df$p_val < 0.05
#selected_df$sig_p_amp <- selected_df$p_amp < 0.05
#selected_df$sig_p_spec <- selected_df$p_spec < 0.05
#selected_df$sig_p_spec_fdr <- selected_df$p_spec_fdr < 0.1
#selected_df$sig_p_time_fdr <- selected_df$p_time_fdr < 0.1

sum(selected_df$pseudo_pval_shuffled < 0.05)
#sum(selected_df$p_time_reject == "True")
sum(selected_df$pseudo_pval_iaaft < 0.05)
#sum(selected_df$sig_p_spec)
#sum(selected_df$p_spec_reject == "True")
#sum(selected_df$sig_p_spec_fdr)

```
Scatter plots to look at differences between me and Casey's methods
1) r_coef (Casey's output) vs. r_obs (my observed rho value) - would expect these to be very similar because I'm trying to do his lagged regression methodology exactly
2) p_time p values (shuffle time, Casey's strategy) and p_spec p values (IAAFT where power spectrum of j is maintained and amplitude is approximated) - don't necessarily expect these to correlate well 

```{r}
model_rho <- lm(spear_rho_real ~ r_coef, data = selected_df)
model_rho
slope_rho <- coef(model_rho)[2]
intercept_rho <- coef(model_rho)[1]

ggplot(data = selected_df, aes(x = r_coef, y = spear_rho_real, color = pseudo_pval_iaaft_signif)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Casey's Spearman rho", 
    y = "Laurie's Spearman rho", title = "Scatter plot of Casey's rho (r_coef) vs. my rho (spear_rho_real)"
  ) +
  theme_minimal()
```

```{r}

ggplot(data = selected_df, aes(x = mean_spec, y = r_obs, color = sig_p_spec)) +
  geom_point() +
  #geom_smooth(method = "lm") +
  labs(
    x = "Mean permuted Spearman rho (IAAFT spec)", 
    y = "Mean empirical Spearman Coefficient", 
    title = "Empirical Spearman Coefficient vs. Mean Permuted Spearman Coefficient"
  ) +
  theme_minimal()
```

```{r}
#example_output$sig_p_time <- example_output$p_time < 0.05

ggplot(data = example_output, aes(x = mean_rho_shuffled, y = spear_rho_real, color = pseudo_pval_shuffled_signif)) +
  geom_point(size = 0.7) +
  #geom_smooth(method = "lm") +
  labs(
    x = "Mean permuted Spearman rho (time-shuffle method)", 
    y = "Mean empirical Spearman Coefficient", 
    title = "Empirical Spearman Coefficient vs. Mean Permuted Spearman Coefficient"
  ) +
  theme_minimal()
```
```{r}
#example_output$sig_p_spec <- example_output$p_spec < 0.05


ggplot(data = example_output, aes(x = mean_rho_iaaft, y = spear_rho_real, color = pseudo_pval_iaaft_signif)) +
  geom_point(size = 0.7) +
  #geom_smooth(method = "lm") +
  labs(
    x = "Mean permuted Spearman rho (IAAFT power-spectrum-aware)", 
    y = "Mean empirical Spearman Coefficient", 
    title = "Empirical Spearman Coefficient vs. Mean Permuted Spearman Coefficient"
  ) +
  theme_minimal()
```

```{r}

ggplot(data = example_output, aes(x = mean_time, y = r_obs, color = sig_p_time)) +
  geom_point() +
  #geom_smooth(method = "lm") +
  labs(
    x = "Mean permuted Spearman rho (time-shuffle method)", 
    y = "Mean empirical Spearman Coefficient", 
    title = "Empirical Spearman Coefficient vs. Mean Permuted Spearman Coefficient"
  ) +
  theme_minimal()
```

```{r}
ggplot(data = selected_df, aes(x = p.value, y = pseudo_pval_shuffled)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(method = "spearman")
  labs(
    x = "Casey's p", 
    y = "Laurie's p", title = "Scatter plot of Casey's p-value from time-shuffle null"
  ) +
  theme_minimal()
```

```{r}
ggplot(data = selected_df, aes(x = pseudo_pval_shuffled, y = pseudo_pval_iaaft)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(method= "spearman") +
  labs(
    x = "p from time-shuffle null", 
    y = "p from IAAFT spec nulll", title = "Scatter plot of p values from time-null vs IAAFT-null"
  ) +
  theme_minimal()
```

Looking into my 10k example output
```{r}
example_output$sig_p_time <- example_output$p_time < 0.05
example_output$sig_p_spec <- example_output$p_spec < 0.05
example_output$sig_p_spec_fdr <- example_output$p_spec_fdr < 0.1
example_output$sig_p_time_fdr <- example_output$p_time_fdr < 0.1

print("OTU pairs with uncorrected p < 0.05 from time shuffle null: ") 
sum(example_output$sig_p_time)
print("OTU pairs with p < 0.05 from time shuffle null AFTER BH-FDR correction: ") 
sum(example_output$p_time_reject == "True")
print("OTU pairs with uncorrected p < 0.05 from iaaft null: ") 
sum(example_output$sig_p_spec)
print("OTU pairs with p < 0.05 from iaaft null AFTER BH-FDR correction: ") 
sum(example_output$p_spec_reject == "True")
print("IF we raise the BH-FDR cutoff to 0.1: ") 
sum(example_output$sig_p_spec_fdr)
sum(example_output$sig_p_time_fdr)


```
## Merge with taxonomy file
```{r}
SILVA_taxonomy <- read_tsv(file = "/Users/laurielyon/Projects/mibi_longitude/data/qiime/taxonomy.tsv")
output_taxonomy_j <- left_join(example_output, SILVA_taxonomy, by = c("OTU_j" = "Feature ID"))


```

```{r}
inner_join_merged_df %>% 
  ggplot(aes(x = p.value, y = pseudo_pval_shuffled)) +
    geom_point()
```



